---
title: "freelance"
author: "mohie eldin"
date: "October, 2019"
output: 
  html_document:
    toc: yes  
    fig_height: 4
    highlight: pygments
    theme: spacelab
---



```{r ,  include=FALSE}
#orders$Year <- substring(orders$`Created at`,1,4) 
#save(orders, file = "orders_data.Rdata")
```


```{r , include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=6) 
library(readr)
library(corrplot)
library(plyr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(tidyr)
library(Hmisc)
library(gridExtra)
library(grDevices)
library(RColorBrewer)
```



***

## Part 1: Data

#### The dataset information is about the orders that have been made from September 2018 to October 2019.
#### It contains originally 64000 observations and 73 columns.
#### I removed 15 column weather because most of them are missing or 99% corelated with another variable. 

### causality :
The data is an `observational` and hence we can not conclude causality from it so we will be able to show `association` between variables and experimental data is needed for further causal analysis.

### Data entry problems :

```{r , echo=FALSE}

#showing the inconsistency in the data entery which make the analysis harder 
head( grep("^[Cc]",unique(orders$`Billing City`),value = TRUE ) , 20)

```

#### There is no unified way in entering the city data and some other columns !

```{r , echo=FALSE}
orders %>% filter(Name == "PS49269") %>% select (Name ,`Financial Status` ,Currency)

```

#### Most of the customer data is written once and the rest of rows are empty !

### So we have to make in the futur a unified way to enter the data through data validation techniques in Excel.
* * *

## Part 2: Research questions

#### 1) What is the  best item category and what is the worst ? 

#### 2) What is the best 5 items and the worst 5 ?

#### 3) Are we able to categorized the customers into A,B and C classes so that we make customized offers to each ? 

#### 4) Do the data reveal some common association between best sellers or between worst sellers or we need more information ?

#### 5) What is the retention rate ?

### Why these are interesting questions ? 

#### Mainly answering these questions will make a deeper understanding to customers and open the doors for a better products and better customer satisfaction.

* * *

## Part 3: Exploratory data analysis

### Q1) What is the best item category and what is the worst ?



```{r,echo=FALSE}
cat("We have ",length(unique(orders$`Lineitem name`)), "different products with 5 categories") 

over_all_sales <- sum(orders$Sales)
over_all_quantity <- sum(orders$`Lineitem quantity`)
categories <- orders %>% group_by(Items_category) %>% summarise(Quantity = sum(`Lineitem quantity`)  , Sales = sum(Sales), Quantity_propotion = round(Quantity/over_all_quantity,2) ,Sales_propotion = round(Sales/over_all_sales,2)
 )                                                    

head( categories %>% arrange(desc(Sales)) %>% as.data.frame())
head(categories %>% arrange(desc(Quantity))%>% as.data.frame())


p1 <- ggplot(data = categories , aes(x = Items_category , y = Sales , fill = Items_category , label = Sales)) +
  geom_bar(stat = "identity") +
  ggtitle("Sales per category") +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_label(aes(fill = Items_category),colour = "white" , fontface = "bold")

p2 <- ggplot(data = categories , aes(x = Items_category , y = Quantity , fill = Items_category , label = Quantity)) +
  geom_bar(stat = "identity") +
  ggtitle("Quantity per category") +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_label(aes(fill = Items_category),colour = "white" , fontface = "bold") 

p1

p2
```


#### We noteced that summer bags soled more than backback; however, their sales are less than backback ! why ? 

#### Also the card holders and slipper are not doing so well in both sales and quantity soled which need more investigation on the reasons behind this.

### Q2) What is the best 5 items and the worst 5 ?


```{r , echo=FALSE}
best_worst_items <- orders %>%  
group_by(`Lineitem name`, Items_category) %>% summarise(quantity_soled = sum(`Lineitem quantity`) , Sales = sum(Sales)) %>% as.data.frame()

by_sales <- head(best_worst_items %>% arrange(desc(Sales)) , n = 5)
by_quantity <-head(best_worst_items %>% arrange(desc(quantity_soled)) , n = 5)

ggplot(data = by_sales , aes(x = reorder(`Lineitem name`,-Sales), y = Sales,fill = by_sales$Items_category  , label =  by_sales$Sales)) + geom_col() + geom_label(colour = "white" , fontface = "bold") +ggtitle(" top 5 best sellers") +
  theme(plot.title = element_text(hjust = 0.5))+ xlab("Item name") 

ggplot(data = by_quantity , aes(x = reorder(`Lineitem name`,-quantity_soled), y = quantity_soled,fill = by_sales$Items_category  , label =  by_quantity$quantity_soled)) + geom_col() + geom_label(colour = "white" , fontface = "bold") +ggtitle(" top 5 best quantity soled") +
  theme(plot.title = element_text(hjust = 0.5)) + xlab("Item name")

```



```{r, echo=FALSE}

worst_by_sales <- tail(best_worst_items %>% filter(`Lineitem name` != "Unknown") %>% arrange(desc(Sales)) , n = 5)

ggplot(data = worst_by_sales , aes(x = reorder(`Lineitem name`,-Sales), y = Sales,fill = worst_by_sales$Items_category  , label =  worst_by_sales$Sales)) + geom_col() + geom_label(colour = "white" , fontface = "bold") +ggtitle(" worst 5 items in sales") +
  theme(plot.title = element_text(hjust = 0.5)) + xlab("Item name")  

```

#### Interestingly the worst five items in sales are not slippers or card holders as expected but the five are summer bags !

```{r, echo=FALSE}
soled_one <- orders %>% group_by(`Lineitem name`,Items_category) %>% summarise(quantity = sum(`Lineitem quantity`)) %>% filter(quantity <= 3)

 freq <- soled_one %>% 
  group_by(Items_category) %>% 
    tally()
 
slices = freq$n
lbls <-  freq$Items_category
pct <- round(slices/sum(slices)*100)
lbls <- paste(lbls, pct) # add percents to labels
lbls <- paste(lbls,"%",sep="") # ad % to labels
pie(slices,labels = lbls, col=c("steelblue" , "purple","yellow","orangered","blue"),
   main="percentage of items who are soled only once in each category")
```

#### Although summer bags are in the best sellers and most demanded items but it also in the worst !

#### Actually we can turn this into a good thing, what about comparing the two groups in terms of the designs, marketing companies, time period and prices !

### Q3) Are we able to categorized the customers into A,B and C classes so that we make customized offers to each ?

```{r , echo=FALSE}
#the propotions of prices 

customer_categories <- orders %>%  filter(`Lineitem price` > 0 , !is.na(Email)) %>% 
   group_by(Email) %>%
    summarise(sales = sum(Sales) , quantity = sum(`Lineitem quantity`) )
sales_classes <- table(cut2(customer_categories$sales , cuts = c(500,1000,359667)))

f=function(x){ifelse(x>0 & x<501,"C",ifelse(x>500 & x <1001,"B","A"))}
customer_categories$sales_category <- sapply(customer_categories$sales , f)
categories_freq <- customer_categories %>% count(sales_category)
sales_classes
```



```{r, echo=FALSE}
slices <- categories_freq$n
lbls <- categories_freq$sales_category
pct <- round(slices/sum(slices)*100)
lbls <- paste(lbls, pct) # add percents to labels
lbls <- paste(lbls,"%",sep="") # ad % to labels

cols <- brewer.pal(3, "BuPu")
pal <- colorRampPalette(cols) 
pie(slices,labels = lbls, col=pal(3),
   main="Propotions of customer categories based on sales")

```


```{r,echo=FALSE}
quantity_classes <- table(cut2(customer_categories$quantity , cuts = c(1,2,3)))
f=function(x){ifelse(x == 1 ,"C",ifelse(x == 2 ,"B","A"))}
customer_categories$quantity_category <- sapply(customer_categories$quantity , f)
categories_freq <- customer_categories %>% count(quantity_category)
quantity_classes
```



```{r,echo=FALSE}
slices <- categories_freq$n
lbls <- categories_freq$quantity_category
pct <- round(slices/sum(slices)*100)
lbls <- paste(lbls, pct) # add percents to labels
lbls <- paste(lbls,"%",sep="") # ad % to labels

cols <- brewer.pal(3, "BuPu")
pal <- colorRampPalette(cols) 
pie(slices,labels = lbls, col=pal(3),
   main="Propotions of customer categories based on quantity they ordered") 

```


```{r, echo=FALSE}
customer_categories$SQ_category <- paste(customer_categories$sales_category , customer_categories$quantity_category)
```

```{r}
customer_categories %>% filter(Email == "alia.samir@msa.edu.eg") %>% as.data.frame()
```

#### Now we have a data frame that contains every thing that we need to know about each class, here is some information about calss A for example.

```{r, echo=FALSE}
class_A_email <- customer_categories %>% filter(sales_category == "A") %>% select(Email)
class_A_email <- class_A_email$Email

class_A_info <- orders %>% filter(Email %in% class_A_email & `Lineitem price` > 0) %>% select(Email , `Lineitem name` , Items_category,Sales,`Lineitem quantity`) 

make_sure <-  class_A_info %>% group_by(Email) %>% summarise(total = sum(Sales)) 

head(class_A_info %>% group_by(`Lineitem name`,Items_category)%>% summarise(quantity_soled = sum(`Lineitem quantity`) )) #5od da screenshot a
```

#### So now we have a deeper insights about what each class had ordered most and we can use this for customized offers ! 


### Q4) Do the data reveal some common association between best sellers or between worst sellers or we need more information ?

```{r , echo=FALSE}
orders %>% filter(`Lineitem price` < 0) %>% group_by(`Lineitem name`,Items_category) %>% summarise(quantity = sum(`Lineitem quantity`) , sales = sum(Sales)) %>% arrange((sales))

orders %>% filter(`Lineitem name` == "15' Cafe x Pink Laptop Bag" & `Lineitem price` > 0) %>% group_by(`Lineitem name`)%>% summarise(sales = sum(Sales) , quantity = sum(`Lineitem quantity`))

```

#### Actually it's wired that the most refunded items are doing well in general so we need to know the reasons of the refunds which may help.

#### Remember the question of why the summer bags hve sales  less than the backbacks ? because of the high rate of refunds !

#### So whta about a column that states the reasons of refund ? we can then work on deacresing this rate. 





```{r , echo=FALSE}

data = orders %>% filter(`Lineitem price` > 0 ) %>%
        group_by(`Lineitem price`,Items_category)  %>%
         summarise(quantity_orderd = sum(`Lineitem quantity`)) %>%
          filter(quantity_orderd < 3000)
price_classes <- table(cut2(data$`Lineitem price` , cuts = c(0,400,600)))

ggplot(data = data , aes(x = data$`Lineitem price`, y = data$quantity_orderd, color = data$Items_category)) + geom_jitter()+ facet_wrap(. ~ data$Items_category)+ geom_smooth(method = "lm" , se = FALSE)+xlab("Price")+ylab("Quantity") + ggtitle("Quantity orderd for each category regarding the prices")
 

```


```{r, echo=FALSE , message=FALSE}

numericVars <- which(sapply(orders , is.numeric))

Cor <- cor(orders[,numericVars], use="pairwise.complete.obs") #correlations of all numeric variables

#sort on decreasing correlations with SalePrice
CorSorted <- as.matrix(sort(Cor[,'Sales'], decreasing = TRUE))
 #select only high corelations with our respone variable
CorHigh <- names(which(apply(CorSorted, 1, function(x) abs(x)>0.5)))
#work on the rows --> margin = 1
#apply return all the column names and True or False so which include the true only with thier names and finally names keeps the names only.
corrplot.mixed( Cor[CorHigh, CorHigh], tl.col="black", tl.pos = "lt" )
```

#### Numerical variables are not helping a lot as expected as the date column is the most important one for a time serise analysis. 

#### we may need to collect more data that can help us estimate items success of filure such as survies, design categories and more. 


### Q5) What is the retention rate ?

#### What is the number of customers in 19 and 18 ?
```{r , echo=FALSE}

#The number of customers :
orders <- orders[which(!is.na(orders$Email)),]
n_18 <- orders %>% filter(Year < "2019") %>% select(Email) %>% unique %>% nrow() 
n_19 <- length(unique(orders$Email))
cat("we had",n_18, "customers in 2018")
cat("Now there are " ,n_19  , "customers in 2019" )
years = c("2018","2019")
number_of_customers = c(n_18,n_19)
x <- data.frame(years,number_of_customers)

ggplot(x , aes(years,number_of_customers,fill = years,label = number_of_customers)) +
  geom_bar(stat = "identity") +
  ggtitle("Number of customer in 2018 and in 2019") +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_label(aes(fill = years),colour = "white" , fontface = "bold") 



  
```

#### This is a huge improvment for increasing the number of customers. 

#### So out of the 25,000 that we have now how many ordered more than once ?
```{r , echo=FALSE}

retention <- orders %>% select(Email,`Created at`) %>%
               group_by(Email) %>%
                summarise(number_of_orders = length(unique(`Created at`))) %>% 
                 filter(number_of_orders > 1)

people_ordered_more_than_once <- nrow(retention)
prop_of_people_ordered_more_than_once = people_ordered_more_than_once/ length(unique(orders$Email))

prop_of_people_ordered_more_than_once <- round(prop_of_people_ordered_more_than_once , 2)

cat("There are " , people_ordered_more_than_once, "bought more than once which is only" ,prop_of_people_ordered_more_than_once *100,"%", "of the customers and we need to know why the rest ",100-prop_of_people_ordered_more_than_once*100,"% didn't order again !"  )

slices <- c(people_ordered_more_than_once, (length(unique(orders$Email))-people_ordered_more_than_once) )
lbls <- c( "Ordered more than once","Ordered once")
pct <- round(slices/sum(slices)*100)
lbls <- paste(lbls, pct) # add percents to labels
lbls <- paste(lbls,"%",sep="") # ad % to labels
pie(slices,labels = lbls, col=c("steelblue" , "orangered"),
   main="Propotions of customers who ordered once and who ordered more than once")

```


#### We can call them, take a feedback and hopefully to make them order again.

## Part 4 : What's next ? 

#### In the future :

#### 1) The data problems has to be handled.
#### 2) More information need to be collected.
#### 3) This will open the door for a better understanding to customers and items and hence machine learning model will be bulit to predict the customer behaviour which is very interesting and promising !





